Reasoning Graphically about Quantum Computation

### 2008-02-24 papers/reasoning-graphically/paper.tex
some changes
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1361 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

\begin{center}
  \scalebox{1.0}{\includegraphics{images/spider_lhs.eps}}
\end{center}

@@

In graph patterns, variable nodes generalise the concept of a
half-edge. A graph with variable nodes is given the following
interpretation in the underyling semantics of our concrete graphs:

@@

instantiation of a variable node:

\begin{itemize}
\item let the set of node kinds be denoted $K$
\item given a graph $G$, containing a variable node $x$, where for a
  given non-vaiable node $n$ it's type is written $T_n$, the concrete
  graph is:

$$
\{G[x/]  | x \in V \}
$$

\item

\end{itemize}

By enriching graphs with variable-nodes and banged-boxes, to form
graph-patterns, we have lost the compositionality principle from the
underlying concrete graphs. However, this can be restored and shown to
correspond to the underyling compositionality principle.

Type Expressions: We introduce a new form of type for a pattern
graph. In particular,

>>>>

\begin{center}
  \scalebox{1.0}{\includegraphics{images/spider_lhs}}
\end{center}

@@

In graph patterns, variable nodes generalise the concept of a
half-edge. Variable nodes, unlike boundary-nodes in the concrete
graph, are not classified into domain and co-domain. The intuition of
a variable node is that it can be replaced by any concrete node. Thus,
unlike nodes in conrete a graph, variable node's do not define an
ordering on the edges that are incident. Instead, the edges that touch
a variable node are simply a set.

The type of a variable nodes is thus defined as a multiset rather than
a tensor. Any tensor type can be lifted into a multiset type in the
obvious manner, which we denote by the operation {\tt
  mset\_of\_tensor}. For example, {\tt mset\_of\_tensor($a \tensor b
  \tensor a$) = $\{(a,2), (b,1)\}$}. We say that a tensor-type, $T$,
is an instance of a multiset type, $M$, when it's lifting into a
multiset is a super-set of $M$: $ M \subseteq
\mathtt{mset\_of\_tensor}(T) \Longrightarrow T \in M$. The intuition
here is that an instance of variable node is allowed to have as many
extra edges as it likes as long as it at least has the ones mentioned
in the original graph. A variable node, $a$ is be said to an instance
of another variable node $b$ when the multiset represnting the type of
$a$ is a superset of the multiset for the type of $b$. Thus, with
regard to the underlying semantics, $a$ being a subtype of $b$ means
that the set of instances of $a$ are a subset of the instances of $b$.

\begin{figure}[t]
%  \scalebox{1.0}{\includegraphics{images/node-var-instance.eps}}
*** Add figure *** 
\label{node-variable-instances-fig}\caption{The concrete graphs $a$,
  $b$, and $c$ are instances of $x$, a graph with variables nodes.}
\end{figure}

A graph with variable nodes is given a formal semantics by being
interpretated as a set of concrete graphs. We refer members of
interpretation set as instances. Formally, an instance of a graph with
variable-nodes is the graph where every variable node has been
replaced by some instance of it. For example, in
Figure~\ref{node-variable-instances-fig}, the graphs $a$, $b$ and $c$
are instances of the graph $x$. Observe that when a variable-node has
degree one, then it can be instantiated as a boundary edge of either
the domain or co-domain.

In this section, we define a compositonality principle for graphs
patterns, that is graphs with variable-nodes and !-boxes. This is
related to compositionality of the underlying concrete graphs by the
following instantiation homomorphism:

By enriching graphs with to form
graph-patterns, we need to define compositionality principle from the
underlying concrete graphs. In this section, we restore
compositionality and show its correspondence to the underyling
principle in concrete graphs.

@@

Type Expressions: we introduce a new form of type for a pattern
graph. In particular, a variable node with
